Stack Safety is a Security Property

### 2021-04-12 tex/main.tex
comments and word-smithing on 2
comments and word-smithing and diagram-smithing on 3

Our initial description of these properties assumes a single simple stack with no sharing between callers and
callees: all parameters and return values are passed in registers.  However,
we can refine it to allow passing of scalar stack data, granting a callee
access to the caller's data. (We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
  \apt{N.B.Doing call-by-reference properly requires that the passed variable is accessible
    by all descendents, not just immediate callee, which is doable with depth tags but not
    obviously with per-activation tags }}

@@

\section{Stack Safety, Intuitively}
\label{sec:running-example}
\colorlet{lgray}{gray!40}
\colorlet{lred}{red!40}
\colorlet{lblue}{blue!20}

@@

int f() {
10:     $SP-2 = 0;
11[R]:  return 5;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 1} &
\memoryaddrs{12em}
\memory{4}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 2} &
\memoryaddrs{12em}
\memory{4}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 10} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}
\end{subfigure}
\label{fig:int1}
\caption{Integrity Violation}
\end{figure}

@@

\begin{itemize}
\item Examples that appear to violate stack safety, violate at least one of our properties
\item We do not find desireable behaviors that appear stack-safe yet violate our properties
\item Our properties are testably implemented by stack safety micro-policies from the
  literature, and the micro-policies are not significantly more conservative
  (i.e., policy failstops usually correspond to imminent property violations)
\end{itemize}

Our examples are written in C-like syntax, labeled with line numbers for reference,
but to reflect the increased power of the underlying assembly code, we allow code
to access the stack pointer directly, written {\tt \$SP}. The layout of the stack
is given to the right, with a box representing a word of memory, and variables stored
in declaration order from left to right. The stack grows upward in memory and to the
right on paper.

@@

Example \ref{fig:int1} shows an integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

The black rectangle represents the return address. In a typical calling convention
this will have been passed in a register during the call itself, and {\tt f} stores
it in the stack during its preamble, so it is not itself sealed.

@@

int f() {
10:     int y,z;
11:     y = 5;
12:     z = $SP-2;
13[R]:  return y+z;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l | l}
  {\tt 1} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{0}
    \MemoryLabel{-14em}{0.75em}{0}
    \MemoryLabel{-10em}{0.75em}{0}
    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 10} &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 11} &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{42}
  \vspace{.5em}
  &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
  \vspace{.5em}
\end{tabular}

@@

\label{fig:conf1}
\end{subfigure}
\caption{A Confidentiality Violation}
\end{figure}

Meanwhile confidentiality intuitively states that a callee is unaware
of the context of its call (except for its arguments and any global variables).
We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
security. Noninterference describes knowledge: it states that for some system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt x}. The trace
shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
state with arbitrary values \(v_0\), etc., in place of the values that should
be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt y}, has
changed to 5 in both the primary trace and the variant. The primary and variant agree on
the location that changed, so confidentiality was not violated. But after line 13,
the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
in the variant trace it might have any value \(v_2\). Say \(v_2 = 43\) -- then the variant
has behaved differently than the primary, violating confidentiality.

This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. Any system components
whose values update will be identical between the primary and variant traces. This includes not
just memory, but registers as well, so that returning a secret is a violation. And when we
extend the model with a notion of external output, it will include any values that are
outputted.

We alternatively could imagine a stepwise form of confidentiality in which we explicitly
track which addresses may be read (i.e., which have been initialized), and at each step
guarantee that the step alters state identically to a step from a variant state. Then we
wouldn't need to divide up the execution by calls. But we can see in this example that taking
a single variant at the start of the call effectively captures initialization: when {\tt y} is
initialized, the primary and the variant agree on its value, so it can now be safely read.
This treatment is roughly in line with more standard uses of noninterference, which typically
vary states at the starts of whole program runs.

@@

int g() {
20:     int z;
21:     if(z == 5)
22[R]:    return 100;
23:     else
24[R]:    return 10;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  \circled{\tt 01} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{0}
    \MemoryLabel{-14em}{0.75em}{0}
    \MemoryLabel{-10em}{0.75em}{0}
    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  \circled{\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  \circled{\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  \circled{\tt 02} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  \circled{\tt 21} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
  \\ &
  \circled{\bf 22}
  ~$\cdots$
  &
  \circled{\bf 24}
  ~$\cdots$
  \\
\end{tabular}
\end{center}

@@

It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt y}
and {\tt z}. In figure \ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
{\tt y} without even directly using its ability to read the stack pointer. In
this case, {\tt g} branches on that value, which is a different way of
violating lockstep confidentiality. Technically, this violation
occurs because the program counter is also a component of the state that must change
identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
the program counters will differ, breaking the lockstep.

@@

\begin{figure}
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int stash = 0;

@@

void f(int &a) {
10:    if(!stash)
11:      stash = &a;
12:    *stash = 5;
13[R]: return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 11} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 11} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}

\label{fig:passing}
\end{figure}

Up til now we have ignored function arguments, since our simplest model places them in
registers. This is an oversimplication, which we will now correct, starting with the more
interesting case: call-by-reference. There are important stack safety concerns
when passing arguments by reference, as seen in example \ref{fig:passing}. In this example,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying the memory to be
passed, and our stack model marks it in magenta. Then the caller's frame that are not
passed are sealed, as usual. Just as we marked sealed memory with the depth of its owner,
we do the same to passed memory. We keep track of the current depth of the active function
next to the stack pointer in the diagram. A passed address may only be accessed when the
current depth is one higher than the one it is sealed with.

@@

Passed addresses are also excluded from confidentiality: at the start of each call, we vary
only stack addresses that are not passed. This is important, as we use the passing label
for those arguments that are passed by value on the stack, as well. We might instead
treat them as unsealed, but then we would treat them as confidential like the rest of the stack.

We do not model permanent sharing, in which a variable's address is taken and it might be
passed far afield before use, because this situation is uninteresting and hard to motivate
as an instance of stack safety. We would expect to treat such sharing as completely
unprotected, and perhaps attach a separate memory safety policy to add some protection
to such variables.

>>>>

Our initial description of these properties assumes a single simple stack with no sharing between callers and
callees: all parameters and return values are passed in registers.  However,
we can refine it to allow passing of scalar stack data, granting a callee
access to the caller's data. (We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
    by all descendents, not just immediate callee, which is doable with depth tags but not
    obviously with per-activation tags }}

@@

\section{Stack Safety, Intuitively}
\apt{section title is a bit sad since we're trying to convert intuition to formal spec!}
\label{sec:running-example}
\colorlet{lgray}{gray!40}
\colorlet{lred}{red!40}
\colorlet{lblue}{blue!20}

@@

int f() {
10:     *($SP-2) = 0;
11[R]:  return 5;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 1} &
\memoryaddrs{12em}
\memory{4}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 2} &
\memoryaddrs{12em}
\memory{4}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 10} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An Integrity Violation}
\label{fig:int1}
\end{figure}

@@

\begin{itemize}
\item examples that appear to violate stack safety violate at least one of our properties;
\item we do not find desireable behaviors that appear stack-safe yet violate our properties; and
\item our properties are testably implemented by stack safety micro-policies from the
  literature, and the micro-policies are not significantly more conservative
  (i.e., policy failstops usually correspond to imminent property violations).
\end{itemize}

Our examples are written in C-like syntax, labeled with line numbers for reference,
but to reflect the increased power of the underlying assembly code, we allow code
to read the stack pointer ({\tt \$SP}) directly and use
it to form pointers.\apt{I've added some strategic *'s}
Call and return annotations are marked {\tt[C]} and {\tt[R]}
respectively. Local variables are assumed to be stored in the stack frame,
in declaration order. Unless otherwise stated, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \$SP} points to the first unused word above the top of the stack. 
At the right of each example, we show the layout of stack memory just before
the specified line number is executed; each box represents a memory word and
addresses increase to the right.

@@

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

The black rectangle represents the return address. In a typical RISC calling convention
this will have been stored in a register by call itself, and {\tt f} stores
it in the stack during its preamble, so it is not itself sealed.

@@

int f() {
10:     int y,z;
11:     y = 5;
12:     z = *($SP-4);
13[R]:  return y+z;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l | l}
  {\tt 1} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{0}
    \MemoryLabel{-14em}{0.75em}{0}
    \MemoryLabel{-10em}{0.75em}{0}
    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 10} &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 11} &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{42}
  \vspace{.5em}
  &
  \memoryaddrs{20em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
  \vspace{.5em}
\end{tabular}

@@

\end{subfigure}
\caption{A Confidentiality Violation}
\label{fig:conf1}
\end{figure}

Meanwhile confidentiality intuitively states that a callee is unaware
of the context of its call (except for its arguments and any global variables).
We state this in a style similar to {\em noninterference}~\citep{??}.
Noninterference describes knowledge: it states that for some system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

In Figure~\ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt x}. The trace
shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
state with arbitrary values \(v_0\), etc., in place of the values that should
be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt y}, has
changed to 5 in both the primary trace and the variant. The primary and variant agree on
the value in the changed location, so confidentiality was not violated. But after line 13,
the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
in the variant trace it might have any value \(v_2\). If \(v_2 \neq 42\) then the variant
has behaved differently than the primary, violating confidentiality.

This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. Any system components\apt{first use of this?}
whose values change will be identical between the primary and variant traces. This includes not
just memory, but registers as well, so that returning a secret is a violation. And when we
extend the model with a notion of external output, it\apt{what?}  will include any values that
emitted.

We alternatively could imagine a stepwise form of confidentiality in which we explicitly
track which addresses may be read (i.e., which have been initialized), and at each step
guarantee that the step alters state identically to a step from a variant state. Then we
wouldn't need to divide up the execution by calls. But we can see in this example that taking
a single variant at the start of the call effectively captures initialization: when {\tt y} is
initialized, the primary and the variant agree on its value, so it can now be safely read.
This treatment is roughly in line with more standard uses of noninterference, which typically
vary states at the starts of whole program runs.\apt{This is all a bit cryptic.}

@@

int g() {
20:     int z;
21:     if(z == 5)
22[R]:    return 100;
23:     else
24[R]:    return 10;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  \circled{\tt 01} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{0}
    \MemoryLabel{-14em}{0.75em}{0}
    \MemoryLabel{-10em}{0.75em}{0}
    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  \circled{\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  \circled{\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  \circled{\tt 02} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{2}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  \circled{\tt 21} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
  \\ &
  \circled{\bf 22}
  ~$\cdots$
  &
  \circled{\bf 24}
  ~$\cdots$
  \\
\end{tabular}
\end{center}

@@

It may be unclear in figure \ref{fig:conf1} why we vary the initial values of {\tt y}
and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
temporal stack safety.
In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
{\tt y} without even directly reading the stack pointer. In
this case, {\tt g} branches on that value, which is a different way of
violating lockstep confidentiality. Technically, this violation
occurs because the program counter is also a component of the state that must change
identically between the primary and variant traces. In a case where \(v_1 \not = 5\),
the program counters will differ, breaking the lockstep.

@@

\begin{figure}
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int *stash = 0;

@@

void f(int &a) {
10:    if(!stash)
11:      stash = a;
12:    *stash = 5;
13[R]: return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 13} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 13} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}
\caption{A violation with pass-by-reference}
\label{fig:passing}
\end{figure}

To this point we have ignored function arguments, since our simplest model places them in
registers. This is an oversimplification, which we will now correct, starting with the more
interesting case: pass-by-reference.
\apt{This is a nice and clear example, but I think the ordering here is confusing. Pass-by-reference and passing arguments on the stack are
  entirely separable issues. I think it would be easiest to explain the latter first, as the
  simplest example of passed arguments.}
There are important stack safety concerns
when passing arguments by reference, as seen in Figure~\ref{fig:passing}. In this example,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying a memory location to be
passed, and our diagram displays it in magenta. The caller's frame locations that are not
passed are sealed as usual. Just as we marked sealed memory with the depth of its owner,
we do the same for passed memory. We keep track of the current depth of the active function
next to the stack pointer in the diagram.\apt{Not seeing that}  A passed address may only be accessed when the
current depth is one higher than the one it is sealed with. \apt{This is not good enough!
  The callee can in turn pass the same argument by reference to its callee...}

@@

Passed addresses are also excluded from confidentiality: at the start of each call, we vary
only stack addresses that are not passed. This is important, as we also use the passing label
for those arguments that are passed by value on the stack. \apt{this is hidden like an afterthought...}
We might instead
treat them as unsealed, but then we would treat them as confidential like the rest of the stack.\apt{unclear.
  would this work or not?}

We do not model permanent sharing, in which a variable's address is taken and it might be
passed far afield before use, because this situation is uninteresting and hard to motivate
as an instance of stack safety. We would expect to treat such sharing as completely
unprotected, and perhaps attach a separate memory safety policy to add some protection
to such variables.\apt{Neither clear nor convincing! Taking explicit addresses of stack
  variables is very interesting to many reviewers!}
